LabMate: supporting types for MATLAB

### 2024-03-26 writing/IMEKOWC2024/LabMate/paper.tex
[ writing ] fantasy part of Section 2

%%%% Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[colorlinks=true]{hyperref}
\usepackage{microtype}
\usepackage{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\remph}{\emph}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

>>>>

%%%% Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[colorlinks=true]{hyperref}
\usepackage{microtype}
\usepackage{proof}
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\newcommand{\remph}{\emph}

\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Consider the problem of resistance calibration: \todo{explain here}.

% $t_i$ time ($\mathsf{T}$; in seconds); $v_i$ voltage ($\mathsf{ML}^2\mathsf{T}^{-3}\mathsf{I}^{-1}$; in volt)

\[
\begin{pmatrix*}
  t_1 & 1 & -1 & 0 & 0 \\
  t_2 & 1 & -1 & 0 & 0 \\
  t_3 & 1 & -1 & 0 & 0 \\
  t_4 & 1 & -1 & 0 & 0 \\
  t_5 & 1 & -1 & 0 & 0 \\
  t_6 & 0 & 0 & 1 & -1 \\
  t_7 & 0 & 0 & 1 & -1 \\
  t_8 & 0 & 0 & 1 & -1 \\
  t_9 & 0 & 0 & 1 & -1 \\
\end{pmatrix*}
\times
\begin{pmatrix}
  m \\
  c \\
  a \\
  c_{\text{offset}} \\
  a_{\text{offset}}
\end{pmatrix}
=
\begin{pmatrix}
  v_1 \\
  v_2 \\
  v_3 \\
  v_4 \\
  v_5 \\
  v_6 \\
  v_7 \\
  v_8 \\
  v_9
\end{pmatrix}
\]

This can be converted into a MATLAB problem with LabMate annotations as follows:
\begin{verbatim}
%> A :: [ 9 x 5 ] double
A = [ t1 1 -1 0 0
      t2 1 -1 0 0
      t3 1 -1 0 0
      t4 1 -1 0 0
      t5 1 -1 0 0
      t6 0 0 1 -1
      t7 0 0 1 -1
      t8 0 0 1 -1
      t9 0 0 1 -1 ]

%> B :: [ 5 x 1 ] double
B = [ m c a cOffset aOffset ]

C = A * B

%> typeof C
\end{verbatim}
%
\todo{Is computing C from A and B what we want?}
The lines starting with `\texttt{\%>}' are interpreted as ordinary comments by MATLAB, but as \remph{input directives} by LabMate. LabMate acts by reading the source code and outputting it again, but possibly with some additional \remph{responses} to the directives added.

Did you notice the error in the above code? The matrix \texttt{B} should be of size $5 \times 1$, but is erroneously of size $1 \times 5$ instead. LabMate detects this error and produces the following as output, where we have suppressed the same definition of \texttt{A} as above:
\begin{verbatim}
%> A :: [ 9 x 5 ] double
A = ...

%> B :: [ 5 x 1 ] double
B = [ m c a cOffset aOffset ]

C = A * B

%> typeof C
%< the type of C is quite a puzzle; unsolved constraints 5 =? 1, 1 =? 5
\end{verbatim}
%
That is, LabMate says that for \texttt{C} to have a sensible type, we need the (impossible) constraints $5 = 1$ and $1 = 5$ to be true, arising from the fact that we have declared \texttt{B} to be a $5 \times 1$ matrix, but in fact it is a $1 \times 5$ matrix. Note that the original source code has been retained, but a response has been inserted to our \texttt{typeof} query. If we change \texttt{B} to
\begin{verbatim}
B = [ m; c; a; cOffset; aOffset ]
\end{verbatim}
instead, then LabMate responds
\begin{verbatim}
%< C :: [ 9 x 1 ] double
\end{verbatim}
This way, LabMate can detect matrix size errors at development time, rather than at runtime.

@@

GravCalc example again: $t_i$ time ($\mathsf{T}$; in seconds); $v_i$ voltage ($\mathsf{ML}^2\mathsf{T}^{-3}\mathsf{I}^{-1}$; in volt)
